---
title: "Regression modelling"
subtitle: "With (life) insurance data"
description: | 
  This article sets out a few practical recipes for modelling with (life) insurance data. Insurance events are typically of low probability - these recipes consider some of   the limitations of "small data" model fitting (where the observations of interest are sparse) and other topics for insurance like comparisons to standard tables. 
author:
  - name: Pat Reen 
    url: https://www.linkedin.com/in/patrick-reen/
categories:
  - statistics
  - regression
  - insurance
theme: main.css 
preview: img/prev_regression.png
date: 2021-10-14
draft: false
output:
  distill::distill_article: 
    toc: true
    toc_float: true
    self_contained: true
    code_folding: true
bibliography: references.bib 
link-citations: yes
---


# Overview 
## Background 
Here I've pulled together a few simple regression tools using (life) insurance data. See earlier article on data transformations/ preparation. Insurance events are typically of low probability - these recipes consider some of the limitations of "small data" model fitting (where the observations of interest are sparse) and other topics for insurance like comparisons to standard tables. Themes covered:

* Grouped vs ungrouped data;
* Choice of: response distribution, link (and offsets), explanatory variables;
* Modelling variance to industry/ reference (A/E or A - E);
* Model selection: stepwise regression, likelihood tests, model evaluation;
* Predictions, confidence intervals and visualisations.
  
See link above to GitHub repository which has the detailed code.

## Libraries
A list of packages used in the recipes.

```{r Setup, message=FALSE, warning=FALSE, results='hide'}
library(rmdformats) # theme for the HTML doc
library(bookdown)   # bibliography formatting
library(kableExtra) # formatting tables
library(scales)     # data formatting  
library(dplyr)      # tidyverse: data manipulation
library(tidyr)      # tidyverse: tidy messy data
library(corrplot)   # correlation matrix visualisation, optional
library(ggplot2)    # tidyverse: graphs
library(pdp)        # tidyverse: arranging plots
library(GGally)     # tidyverse: extension of ggplot2
library(ggthemes)   # tidyverse: additional themes for ggplot, optional
library(plotly)     # graphs, including 3D 
library(caret)      # sampling techniques
library(broom)      # tidyverse: visualising statistical objects
library(pROC)       # visualising ROC curves
library(lmtest)     # tests for regression models

# packages below have some interaction with earlier packages, not always needed
library(arm)        # binned residual plot
library(msme)       # statistical tests, pearson dispersion
library(MASS)       # statistics

```

```{r Setup - remove object if exists, warning = FALSE, echo=FALSE, results = FALSE, message=FALSE, include=FALSE}
ifrm <- function(x, env = globalenv()) 
{
  if(exists(x, envir = env)) 
  {
    rm(list = x, envir = env)
  }
}
```

```{r Detatch MASS, warning = FALSE, echo=FALSE, results = FALSE, include=FALSE}
# MASS package interferes with pairs() and cor() and others
require(arm) # package
detach(package:arm)
detach(package:msme)
detach(package:MASS)
```

## Further reading
A few books and articles of interest:

* [R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/) has everything you need to know to set up an r markdown document like this one.
* [Generalised Linear Models for Insurance Data](https://www.cambridge.org/au/academic/subjects/statistics-probability/statistics-econometrics-finance-and-insurance/generalized-linear-models-insurance-data?format=HB&isbn=9780521879149) is a great book introducing GLMs in the context of insurance, considering problems specific to insurance data.
* [Tidyverse documentation](https://www.tidyverse.org/) full set of documentation for Tidyverse packages (packages for data science) e.g. dplyr for data manipulation; tidyr for tidying up messy data; ggplot for visualisation.

# Model selection 
The sections below provide a refresher on linear and logistic regression; some considerations for insurance data; model selection and testing model fit.

## Splitting data

### Training vs testing data
Split data into training and testing data sets. We will used 75% of the data for training and the rest for testing.

```{r Load data, warning = FALSE, echo=FALSE, results = FALSE, message=FALSE, include=FALSE}
rm(list = ls())
load(file = "data.RData",)
```

```{r Split data}
# Determine the number of rows for training
nrow(df)
# Create a random sample of row IDs
sample_rows <- sample(nrow(df),0.75*nrow(df))
# Create the training dataset
df_train <- df[sample_rows,]
# Create the test dataset
df_test <- df[-sample_rows,]
```

### Class imbalance
Looking at the random sample we have created, we have a significant imbalance between successes and failures. 

```{r Split data - review}
df_train %>% select(inc_count_acc,inc_count_sick) %>% table()
```

Generally, "in insurance applications, the prediction of a claim or no claim on an individual policy is rarely the point of statistical modelling ... The model is useful provided it explains the variability in claims behaviour, as a function of risk factors. " [@GLM_Insurance, p108-109] So, as long as we have sufficient actual claims to justify the level of predictor variables fitted to the model we should be ok.

However, in some cases where it is important to correctly predict the binary categorical response variable, we may need to create a sample that has a roughly equal proportion of classes (of successes and failures) and then fit our model to that data. E.g. where we are looking to accurately predict fraud within banking transactions data.

To do that we need to refine our sampling method

* Down sampling: the majority class is randomly down sampled to be of the same size as the smaller class. 
* Up sampling: rows from the minority class (e.g. claim) are repeatedly sampled over and over until they reaches the same size as the majority class (not claim).
* Hybrid sampling: artificial data points are generated and are systematically added around the minority class.

Showing a method for down sampling below - this is more useful for pure classification models; not that useful for insurance applications.

```{r Split data - class imbalance}
# Determine the number of rows for training
df_train_ds <- downSample(df_train,factor(df_train$inc_count_sick)) 
df_train_ds %>% select(inc_count_sick) %>% table()
```

## Regression 

### Background
Linear regression is a method of modelling the relationship between a response (dependent variable) and one or more explanatory variables (predictors; independent variables). For a data set , the relationship between y, the response/ dependent variables and the vector of x's, the explanatory variables, is linear of the form

$y_{i} = \beta_{0} + \beta_{1}x_{i1} + ... + \beta_{p}x_{ip} + \epsilon_{i} = \mathbf{x}_{i}^t\mathbf{\beta} + \epsilon_{i}$, $i = 1,...,n$

Key assumptions

* linearity - response is some linear combination of regression coefficients and explanatory variables
* constant variance (homoskedastic) - variance of errors does not depend upon explanatory variables
* errors are independent - response variables uncorrelated
* explanatory variables are not perfectly co-linear
* weak exogeneity - predictors treated as fixed / no measurement error.

### Linear vs logistic regression 

* Linear: Continuous response outcome i.e. predicting a continuous response/dependent variable using a set of explanatory variables
* Logistic: Binary response outcome - straight line does not fit the data well. The predicted values are always between 0 and 1. 

For logistic regression the log odds are linear. We can transform to odds ratios by taking the exponential of the coefficients or exp(coef(model)) - this shows the relative change to odds of the response variables, where

*	Odds-ratio = 1, the coefficient has no effect. 
*	Odds-ratio is <1, the coefficient predicts a decreased chance of an event occurring. 
*	Odds-ratio is >1, the coefficient predicts an increased chance of an event occurring.
	
### Logistic regression for claims incidence
A simple regression model is shown below. Interpreting co-efficients and other output:

* Intercept - global intercept and reference intercepts for each group (reference intercept: + contract (y~x); or intercept for each group: y~x-1)
* Slopes - other coefficients, estimates linear coefficient for continuous variable

Other output

* Pr(>|z|)/ p value is the probability coefficient result you're seeing happened due to random variation. Commonly a p-value of .05 or less is significant.
* AIC and likelihood tests, useful for model comparison.
* Residuals sections gives summary stats of the model.
* Call gives the form of the model.
* vcov(model) gives the variance-covariance matrix for fitted model (diagonals are variance and off diagonals are covariance - 0 if all variables are orthogonal).
* coef(model) returns the model coefficients.

```{r Logistic regression - base}
model_1 <- glm(inc_count_sick~age,data=df_train,family="binomial") # use the default link
summary(model_1)
```

Choice of [link function](https://en.wikipedia.org/wiki/Generalized_linear_model): the link function defines the relationship of response variables to mean. It is usually sufficient to use the standard link. Section below shows a model with the link specified (results are the same as the model above).

```{r Logistic regression - link function, results='hide'}
model_2 <- glm(inc_count_sick~age,data=df_train,family=binomial(link="logit")) # specify the link, logit is default for binomial
summary(model_2)
```

Adding more response variables: When deciding on explanatory variables to model, consider the properties of the data (like correlation or colinearity). 

```{r Logistic regression - more variables}
model_3 <- glm(inc_count_sick~age+policy_year+sex+waiting_period,data=df_train,family=binomial(link="logit")) 
summary(model_3)
```

### Linear regression for claims incidence
Choice of response distribution: In the logistic regression example above we considered the Binomial response distribution. Other common count distributions are

* [Poisson](https://en.wikipedia.org/wiki/Poisson_distribution) distribution: mean and variance are equal. 

```{r Response distribution - count densities, poisson}
lower<-qpois(0.001, lambda=5)
upper<-qpois(0.999, lambda=5)
x<-seq(lower,upper,1)

data.frame(x, y=dpois(x, lambda=5)) %>% ggplot(mapping=aes(y=y,x=x)) + geom_col()+
labs(x=NULL, y="Density", title = "Density, poisson [dpois(x,lambda=5)]")+
theme_pander() + scale_color_gdocs()
```

* [Negative Binomial](https://en.wikipedia.org/wiki/Negative_binomial_distribution) distribution: can handle Poission overdispersion (where the variance is bigger than expected).

```{r Response distribution - count densities, negative binomial}
lower<-qnbinom(0.001, size=2, mu=10)
upper<-qnbinom(0.999, size=2, mu=10)
x<-seq(lower,upper,1)

data.frame(x, y=dnbinom(x, size=2, mu=10)) %>% ggplot(mapping=aes(y=y,x=x)) + geom_col()+
labs(x=NULL, y="Density", title = "Density, negative binomial [dnbinom(x, size=2, mu=10)]")+
theme_pander() + scale_color_gdocs()
```

And amount

* [Normal](https://en.wikipedia.org/wiki/Normal_distribution) distribution

```{r Response distribution - amount densities, normal}
x<-seq(-3.5,3.5,0.1)  
data.frame(x,y=dnorm(x, mean=0, sd=1)) %>% ggplot(mapping=aes(y=y,x=x)) + geom_line()+
labs(x=NULL, y="Density", title = "Density, standard normal [dnorm(x, mean=0, sd=1)]")+
theme_pander() + scale_color_gdocs()
```

* [Gamma](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/GammaDist.html) distribution: allows for a fatter tail.

```{r Response distribution - amount densities, gamma}
lower<-qgamma(0.001, shape=5, rate=3)
upper<-qgamma(0.999, shape=5, rate=3)
x<-seq(lower,upper,0.01)

data.frame(x,y=dgamma(x, shape=5, rate=3)) %>% ggplot(mapping=aes(y=y,x=x)) + geom_line()+
labs(x=NULL, y="Density", title = "Density, gamma [dgamma(x,shape=5, rate=3)]")+
theme_pander() + scale_color_gdocs()
```

An example of an amount model, assuming a normal response distribution

```{r Linear regression}
amount_model_3 <- glm(inc_amount_sick~age+sex+waiting_period,data=df_train) 
tidy(amount_model_3)
```

## (Un)grouped data
The models above were based upon ungrouped data, but earlier we noted that data are often grouped. In grouping data, some detail is usually lost e.g. we no longer have any true continuous categorical response variables. In this case, the columns relating to actual events (here claims) are a sum of the individual instance; more at @GLM_Insurance, p49, 105.

The example below compares a model of counts by age and sex on the grouped and shows that the derived parameters are materially similar.

```{r Logistic regression - grouped}
# grouped model
model_4 <- glm(cbind(inc_count_sick,exposure-inc_count_sick)~age+sex,data=df_grp,family=binomial(link="logit")) # cbind gives a matrix of successes and failures
```

Use tidy to visualise the modelled result for grouped data:

```{r Logistic regression - grouped tidy}
tidy(model_4)
```

Modelling using ungrouped data, coefficients materially similar

```{r Logistic regression - ungrouped}
model_5 <- glm(inc_count_sick~age+sex,data=df,family=binomial(link="logit")) 
```

Use tidy to visualise the modelled result for ungrouped data:

```{r Logistic regression - ungrouped tidy}
tidy(model_5)
```

### Offsets
When the data are grouped, it is important to consider the level of exposure in a given group. This can be achieved with an [offset](https://towardsdatascience.com/offsetting-the-model-logic-to-implementation-7e333bc25798) term (@GLM_Insurance, p66-67). This is important in an insurance context where we are often interested in modelling rates of claim. 

```{r Logistic regression - offset}
df_grp_filtered <- df_grp %>% filter(inc_count_tot>1, exposure>10)
model_6 <- glm(inc_amount_tot~age+sex+offset(sum_assured),data=df_grp_filtered) 
tidy(model_6)
model_7 <- glm(inc_amount_tot~age+sex,data=df_grp_filtered) 
tidy(model_7)
```

## Actuals or AvE?
The models we have fitted above are based upon actual claim incidences. We can consider the difference between some expected claim rate (e.g. from a standard table) and the actuals. 

Using the grouped data again, we model A/E with a normal response distribution. The model shows that none of the coeffients are significant indicating no material difference to expected, which is consistent with the data as the observations were derived from the expected probabilities initially.

```{r Logistic regression - grouped A/E}
# grouped model
model_8 <- glm(inc_count_sick/inc_count_sick_exp~age+sex,data=df_grp) 
# use tidy to visualise the modelled result
summary(model_8)
```

## Stepwise regression
This method builds a regression model from a set of candidate predictor variables by entering predictors based on p values, in a stepwise manner until there are no variables left to enter any more. Model may over/ understate the importance of predictors and the direction of stepping (forward or backward) can impact the outcome - so some degree of interpretation is necessary.

```{r Stepwise regression}
# specify a null model with no predictors
null_model_sick <- glm(inc_count_sick ~ 1, data = df_train, family = "binomial")

# specify the full model using all of the potential predictors
full_model_sick <- glm(inc_count_sick ~ cal_year + policy_year + sex + smoker + benefit_period + waiting_period + occupation + poly(age,3) + sum_assured + policy_year*age + policy_year*sum_assured, data = df_train, family = "binomial")

# alternatively, glm(y~ . -x1) fits model using all variables excluding x1

# use a forward stepwise algorithm to build a parsimonious model
step_model_sick <- step(null_model_sick, scope = list(lower = null_model_sick, upper = full_model_sick), direction = "forward")

summary(full_model_sick)
summary(step_model_sick)

```

The form of the final step model is glm(formula = inc_count_sick ~ poly(age, 3) + waiting_period + sex + smoker + sum_assured, family = "binomial", data = df_train) i.e. dropping some of the explanatory variables from the full model.

## Confidence intervals
We can compute the confidence intervals for one or more parameters in a fitted model.

```{r Stepwise regression - confidence intervals, message=FALSE, warning=FALSE}
confint(step_model_sick) # add second argument specifying which parameters we need a confint for
```

## Predictions
predict() is a generic function that can be used to predict results from various model forms. The function takes the form below. For logistic regression, setting prediction type to response produces a probability rather than log odds (which are difficult to interpret).

```{r Predictions}
# predictions
pred_inc_count_sick <- as.data.frame(
  predict(step_model_sick, data= df_train, # model and data
  type="response", # or terms for coefficients
  se.fit = TRUE, # default is false
  interval = "confidence", #default "none", also "prediction"
  level = 0.95
  # ...
  )
)

# add back to data
ifrm("df_train_pred")
pred_inc_count_sick <- rename(pred_inc_count_sick,pred_rate_inc_count_sick=fit,se_rate_inc_count_sick=se.fit)
df_train_pred <- cbind(df_train,pred_inc_count_sick)
```

We can plot the results by age and sex against the crude rates from earlier:

```{r Predictions - graph}
# from earlier, summarise data by age and sex
df_train_pred %>% filter(sex != "u", between(age, 30,60)) %>% group_by(age,sex) %>% 
summarise(total_sick=sum(inc_count_sick),total_acc=sum(inc_count_acc), pred_total_sick=sum(pred_rate_inc_count_sick),exposure=n(),.groups = 'drop') %>% 
mutate(sick_rate = total_sick/exposure,pred_sick_rate = pred_total_sick/exposure, acc_rate = total_acc/exposure) %>%
# used ggplot to graph the results
ggplot(aes(x=age,y=sick_rate,color=sex)) +
# ylim(0,1) +
geom_point() +
# add a modeled line
geom_line(aes(x=age,y=pred_sick_rate)) +
theme_pander() + scale_color_gdocs()
```

### Out of sample predictions
Earlier we split the data into “training” data used to create the model and “test” data which we intended to use for performance validation. Adding predictions to test data below.

```{r Predictions for validation, message=FALSE, warning=FALSE}
# predictions
pred_inc_count_sick <- as.data.frame(
  predict(step_model_sick, data= df_test, # test data
  type="response",
  se.fit = TRUE, 
  interval = "confidence",
  level = 0.95
  # ...
  )
)

# add back to data
ifrm("df_test_pred")
pred_inc_count_sick <- rename(pred_inc_count_sick,pred_rate_inc_count_sick=fit,se_rate_inc_count_sick=se.fit)
df_test_pred <- cbind(df_test,pred_inc_count_sick)

# summary stats for prediction
hist(df_test_pred$pred_rate_inc_count_sick,main = "Histogram of predicted sickness rate",xlab = "Probability of claim")
summary(df_test_pred$pred_rate_inc_count_sick)

```

To translate the predicted probabilities into a vector of claim/no-claim for each policy we could define a claim as occurring if modelled/ predicted probability of claim is greater than some threshold value. More on this later under evaluation techniques. 

```{r Binary predictions}
# add binary prediction based upon a threshold probability
df_test_pred$pred_inc_count_sick <- ifelse(df_test_pred$pred_rate_inc_count_sick>0.003,1,0) # example threshold is ~3rd quartile probability of claim. for balanced data this should be closer to 0.5.
```

# Evaluation 
## Techniques

### AIC
For least squares regression the $R^{2}$ statistic ([coefficient of determination](https://en.wikipedia.org/wiki/Coefficient_of_determination)) measures the proportion of variance in the dependent variable that can be explained by the independent variables. Adjusted $R^{2}$, adjusts for the number of predictors in the model. The adjusted $R^{2}$ increases when the new predictor improves the model more than would be expected by chance. The glm function uses a maximum likelihood estimator which does not minimize the squared error.

AIC stands for [Akaike Information Criteria](https://en.wikipedia.org/wiki/Akaike_information_criterion). $AIC = -2l+2p$ where $l$ is the log likelihood and $p$ are the number of parameters. It is analogous to adjusted $R^{2}$ and is the measure of fit which penalizes model for the number of independent variables. We prefer a model with a lower AIC value.

```{r Re-attatch MASS, warning = FALSE, echo=FALSE, results = FALSE, message=FALSE, include=FALSE}
# MASS package interferes with pairs() and cor(), reload here for stats tests later
library("MASS")
library("arm")
library("msme")
```

The results below show the AIC for model 3 is lower than model 1 and that the final step model has the lowest AIC of those evaluated (preferred).

```{r Evaluation - AIC}
AIC(model_1)
AIC(model_3)
stepAIC(step_model_sick)
```

### Anova

An anova comparison below between model_3 and the step model using a Chi-squared test shows a small p value for the stepped model - indicating that the model is an improvement. F-test can be used on continuous response models.

```{r Evaluation - anova}
anova(model_3,step_model_sick, test="Chisq")
```

### Likelihood ratio test
The [likelihood ratio test](https://en.wikipedia.org/wiki/Likelihood-ratio_test) compares two models based upon log likelihoods; more at @GLM_Insurance, p74.

The test below concludes that the step model is more accurate than the less complex model.

```{r Evaluation - likelihood ratio}
lrtest(model_3,step_model_sick)
```

### Other tests
Other validations tests could be considered like the [Wald test](https://en.wikipedia.org/wiki/Wald_test), [Score test](https://en.wikipedia.org/wiki/Score_test); see @GLM_Insurance, p74-77.

## Residual checks
Residuals/ errors are the observed less fitted values. Traditional residual plots (shown below) are usually a good starting point (we expect to see no trend in the plot of residuals vs fitted values), but are not as informative for logistic regression or for data with a low probability outcome. 

### Standard model plots
Plots, linear regression example:

```{r Evaluation - traditional plots linear reg}
par(mfrow = c(2, 2))
plot(amount_model_3)
```

Logistic regression example:

```{r Evaluation - traditional plots logistic reg}
par(mfrow = c(2, 2))
plot(step_model_sick)
```

### Alternatives
For logistic regression we can try other tools to test the residuals against the model assumptions. GLMs assume that the residuals/ errors are normally distributed. Plotting the density of the residuals gives us:

```{r Evaluation - histogram of residuals}
hist(rstandard(step_model_sick),breaks= c(seq(-1,1,by=0.001),seq(2,5, by=1)),freq=FALSE,main = "Histogram of residuals",xlab = "Residuals")
curve(dnorm, add = TRUE)
```

Focusing the x axis range:

```{r Evaluation - histogram of residuals focussed}
hist(rstandard(step_model_sick),breaks= c(seq(-1,1,by=0.001),seq(2,5, by=1)), xlim = c(-0.5,0.5),freq=FALSE,main = "Histogram of residuals",xlab = "Residuals")
```

A binned residual plot divides data into bins based upon their fitted values, showing the average residuals vs fitted value for each bin [@stats_notes]:

```{r Evaluation - binned residual}
# from the arm package
binnedplot(fitted(step_model_sick), 
           residuals(step_model_sick, type = "response"), 
           nclass = 50, 
           xlab = "Expected Values", 
           ylab = "Average residual", 
           main = "Binned residual plot", 
           cex.pts = 0.8, 
           col.pts = 1, 
           col.int = "gray")
```

Grey lines are 2 se bands (~95%). Apart from a few outliers, most of the residuals are within those bands.

### P-P plots
The P-P (probability–probability) plot is a visualization that plots CDFs of the two distributions (empirical and theoretical) against each other, an unrelated dummy example below. It can be used to assess the residuals for normality.

```{r Evaluation - pp plot}
x <- rnorm(100)
probDist <- pnorm(x)
#create PP plot
plot(ppoints(length(x)), sort(probDist), main = "PP Plot", xlab = "Observed Probability", ylab = "Expected Probability")
#add diagonal line
abline(0,1)
```

### Other tests

[Pearson dispersion test](https://search.r-project.org/CRAN/refmans/msme/html/P__disp.html): This function calculates Pearson Chi2 statistic and the Pearson-based dispersion statistic. Values of the dispersion greater than 1 indicate model overdispersion. Values under 1 indicate under-dispersion.

```{r Evaluation - pearson dispersion}
P__disp(step_model_sick)
```

## Confusion matrix
With a binary prediction from the model loaded within the dataframe (defined earlier), we can compare this to the actual outcomes to determine the validity of the model. In this comparison, the true positive rate is called the Sensitivity. The inverse of the false-positive rate is called the Specificity.

* Sensitivity = TruePositive / (TruePositive + FalseNegative)
* Specificity = TrueNegative / (FalsePositive + TrueNegative)

Where:

* Sensitivity = True Positive Rate
* Specificity = 1 – False Positive Rate

A perfect classification model could have Sensitivity and Specificity close to 1. However, we noted earlier, in insurance applications we are not often interested in in an accurate prediction as to whether a given policy gives rise to a claim. Rather we are interested in understanding the claim rates and how they are explained by the response variables/ risk factors [@GLM_Insurance, p108-109].

A confusion matrix is a tabular representation of Observed vs Predicted values. It helps to quantify the efficiency (or accuracy) of the model.

```{r Detatch MASS - 2, warning = FALSE, echo=FALSE, results = FALSE, include=FALSE}
# MASS package interferes with pairs() and cor() and others
require(arm) # package
detach(package:arm)
detach(package:msme)
detach(package:MASS)
```

```{r Evaluation - confusion matrix}
confusion_matrix <- df_test_pred %>% select(inc_count_sick,pred_inc_count_sick) %>% table()
confusion_matrix
```

```{r Evaluation - confusion matrix - detail}
cat("Accuracy Rate =",(confusion_matrix[1,1]+confusion_matrix[2,2])/sum(confusion_matrix[]),
"; Missclasification Rate =",(confusion_matrix[1,2]+confusion_matrix[2,1])/sum(confusion_matrix[]),
"; True Positive Rate/Sensitivity  =",confusion_matrix[2,2]/sum(confusion_matrix[2,]))

cat("False Positive Rate =",confusion_matrix[1,2]/sum(confusion_matrix[1,]),
"; Specificity =",1-confusion_matrix[1,2]/sum(confusion_matrix[1,]))
```

## ROC/ AUC
The ROC (Receiver Operating Characteristic) curve is a graph with:

* The x-axis showing the False Positive Rate
* The y-axis showing the True Positive Rate

ROC curves start at 0 on the x and y axis and rise to 1. The faster the curve reaches a True Positive Rate of 1, the better the curve generally. A model on the diagonal is only showing a 50/50 chance of correctly guessing the probability of claim.  Area under an ROC curve (AUC) is a measure of the usefulness of a model in general, where a greater area means more useful. AUC is a tool for comparing models (generally, the closer the AUC is to 1, the better, but there are some cases where AUC can be misleading; AUC = 0.5 is a model on the diagonal).

In our binary model, the AUC is not much better than 0.5, indicating a very weak predictive ability. It isn't hugely surprising that our model is not very effective at predicting individual claims. As noted earlier, for insurance applications, we are usually more concerned with predicting the claim rate and how it varies by predictors like age and sex [@GLM_Insurance, p108-109].

```{r Evaluation - ROC, warning = FALSE, message=FALSE}
par(mfrow = c(1,1))
roc = roc(df_test_pred$inc_count_sick, df_test_pred$pred_rate_inc_count_sick, plot = TRUE, print.auc = TRUE)
as.numeric(roc$auc)
coords(roc, "best", ret = "threshold")
```

# References 

<div id="refs"></div>


